---

#Analysis Of Data From Activity Monitoring Device
#Rohit Banerjee
#Date: February 10, 2018


  
    
#Synopsis:
  
With the advancement of technology ,it is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

Here we have used an activity data set to analyse and understand the pattern of steps  of an individual  taken at 5 minute intervals.

```{r}
library(knitr)
library(dplyr)
opts_chunk$set(echo = TRUE)
library(ggplot2)

```
#Loading the Dataset into R
1.1 Reading the dataset into R

1.2 converting the dates into date class variable
```{r}
activitydata <- read.csv2("activity.csv",sep = ",",header = TRUE)
activitydata$date = as.Date(activitydata$date, "%Y-%m-%d")

```
#An overview of the Dataset being used
```{r}
head(activitydata)
```




#Histogram of the total number of steps taken each day
2.1 Copying the original data to another data frame for analysis.

2.2 Converting the dates into factors

2.3 Summarising the daily steps grouped by the dates.
```{r}
testdata <- activitydata
testdata$date <- as.factor(testdata$date)
grouped_dates <- testdata %>% group_by(date) %>% summarise(steps=sum(steps))
grouped_dates$date <-as.Date(grouped_dates$date,"%Y-%m-%d")
head(grouped_dates)
```
2.4 Plotting the histogram for total number of steps taken each day.
```{r}
plot(grouped_dates$steps~grouped_dates$date,type = "h", lwd = 4 , col = "steel blue", xlab = "Dates" , ylab = "Steps" ,main = "Histogram for total number of steps taken each day")
abline(h = max(grouped_dates$steps,na.rm = TRUE))
```



#Mean and median number of steps taken each day
3.1 Mean steps taken

3.2 Median of the steps taken
```{r}
mean_steps <- mean(grouped_dates$steps, na.rm = TRUE)
mean_steps
median_steps <- median(grouped_dates$steps,na.rm = TRUE)
median_steps
```
#Time series plot of the average number of steps taken
4.1 Getting the average of the steps on each interval
```{r}
average_steps <- testdata %>% filter(!is.na(steps)) %>% group_by(interval) %>% summarise(steps = mean(steps)) 
```
4.2 Time series plot of the data
```{r}
ggplot(average_steps,aes(x = interval, y = steps)) + geom_line(color ="steel blue") + labs(title = "Time series plot of the average number of steps taken" , x= " Interval " , y = " Steps Taken ")

```


#The 5-minute interval that, on average, contains the maximum number of steps
```{r}
average_steps[(which.max(average_steps$steps)),]
```


#Code to describe and show a strategy for imputing missing data
6.1 Checking for the steps with Na values

6.2 Calculating the average of steps at 5 minute time intervals

6.3 Replacing the Na values with the corresponding average steps taken. 

```{r}
step_na <- is.na(testdata$steps)
average_steps <- tapply(testdata$steps,testdata$interval,mean,na.rm = TRUE,simplify = TRUE)
testdata$steps[step_na] <-average_steps[as.character(testdata$interval[step_na])] 
head(testdata)
```
#Histogram of the total number of steps taken each day after missing values are imputed
7.1 Grouping the new formed data based on the intervals and weekdays.

7.2 Plotting the data using ggplot after imputing the missing values. 
```{r}
steps_eachday <- testdata %>% group_by(interval) %>% summarise(steps = sum(steps))
plot(steps_eachday$steps~steps_eachday$interval , type = "h" , lwd = 4, xlab = " Interval ", ylab = " Steps ", main = "Histogram of the number of steps after missing values are imputed", col = " steel blue ")
```


#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

8.1 Converting the dates into Weekdays

8.2 Plotting the data segregated into Weekend and Weekdays.

```{r}
library(lubridate)
testdata$date <- as.Date(testdata$date , "%Y-%m-%d")
 testdata$weekdays <- ifelse(weekdays(testdata$date) == "Saturday" | weekdays(testdata$date) == "Sunday" ,"Partydays(Weekend)" , "Sloppydays(Weekdays)")
grouped_weektype <- testdata %>% group_by(interval , weekdays) %>% summarise(steps = mean(steps))
ggplot(grouped_weektype, aes(x=interval, y=steps, color = weekdays)) + geom_line() + facet_wrap(~weekdays, ncol = 1, nrow=2)

```




